---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# padaptR

<!-- badges: start -->
<!-- badges: end -->

padaptR is a tool which makes it easier to work with data downloaded from PADAPT (Pannonian Database Of Plant Traits). They also have a website: https://padapt.eu/en

The original authors of this database are Sonkoly et. al., their manuscript is available at https://doi.org/10.1038/s41597-023-02619-9

## Overview

So far these are the available functions within this package:  
`padapt_query()` Searches the database and applies the filters you need  
`search_species()` Returns all the species names which contains the search string  
`check_typo()` Validate the species names, get recommendations in case of error  
`rand_species() / rand_traits()` Get traits, and species names randomly  
`convert()` Convert standard scientific names into the ones found in PADAPT, also searches
for synonyms from GBIF

## Bug reports:  

If you ever came across a bug, either:  
* open an issue on github  
* contact me on this e-mail address: sutacdominik@gmail.com  

## Installation

You can install the development version of padaptR from [GitHub](https://github.com/dominiksutac/padaptR) with:

``` {r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("dominiksutac/padaptR")
```

```{r}
library(padaptR)
```
## Examples

In the following I am gonna demonstrate how all of these works currently

```{r}
set.seed(42)
rand_species(2)
rand_traits(2)
```

```{r}
#We know there are a few Stipa species but let's inspect them
search_species('Stipa')
```

```{r}
# Let's have Stipa pennata L.
sp = 'Stipa pennata L.'
check_typo(sp)

#Now let's introduce a mistake
sp2 = 'Spita pennata L.'
check_typo(sp2)
```

```{r}
species = c("Achillea collina ","Adonis vernalis","Agropyoron intermedium","Agrostis stolonifera","Agrostis tenuis","Koeleria gracilis" )

# We can specify the output, by default we'll only get the names we managed to match, and the ones
# we did not found a match for
convert(species) 

# We also have an option to inspect what our input has been matched to

convert(species, table = T)

```

```{r}
padapt_query(list_of_species = 'Stipa pennata L.', 
             preset = 'leaf_traits', means = T, just_means = F)
padapt_query(list_of_species = c('Stipa pennata L.', 'Stipa capillata L.'), 
             preset = 'leaf_traits', means = T, just_means = T)
```
